---
title: "Stochastic SIR model using odin"
author: "Thibaut Jombart, Rich FitzJohn"
date: "`r Sys.Date()`"
---

``` {r setup, echo = FALSE, results = "hide"}
lang_output <- function(x, lang) {
  cat(c(sprintf("```%s", lang), x, "```"), sep = "\n")
}
c_output <- function(x) lang_output(x, "cc")
r_output <- function(x) lang_output(x, "r")
plain_output <- function(x) lang_output(x, "plain")
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 5,
  collapse = TRUE)
```

# Basic model

The stochastic equivalent of the previous model can be formulated
in `odin` as follows:

``` {r}
model_file <- "odin_model.R"
```

``` {r echo = FALSE, results = "asis"}
r_output(readLines(model_file))
```


We can use the same workflow as before to run the model, using 10
initial infected individuals (`I_ini = 10`):
``` {r }
sir_generator <- odin::odin(model_file, verbose = FALSE)
sir_generator
x <- sir_generator(I_ini = 10)
```

``` {r fig.cap = "<i>An example of stochastic, discrete-time SIR model</i><br>"}
set.seed(1)
sir_col <- c("#8c8cd9", "#cc0044", "#999966")

res <- x$run(0:100)
par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
matplot(res[, 1], res[, -1], xlab = "Time", ylab = "Number of individuals",
        type = "l", col = sir_col, lty = 1)
legend("topright", lwd = 1, col = sir_col, legend = c("S", "I", "R"), bty = "n")
```


This gives us a single stochastic realisation of the model, which
is of limited interest.



# Generating replicates

Stochastic models become interesting when multiple *replicates*
(i.e. independent stochastic realisations of the model) can be generated.
*odin* implements a straightforward interface for this, which we illustrate
below. 

First, let's have a look at the arguments of the `run` function of our new object:
```{r}
args(x$run)
```

`replicate` can take an `integer` indicating the number of independent
realisations of the model we are after. Let's start with 200:
```{r}
res_200 <- x$run(0:100, replicate = 200)
dim(res_200)
```

`res_200` is now a 3-dimensional array storing results for all time steps,
compartments and replicates. The function `x$transform_variables` will be very
handy to reformat these results into something we can plot and analyse:

```{r}
res_200 <- x$transform_variables(res_200)
class(res_200)
length(res_200)
names(res_200)
dim(res_200[[1]])
```

`res_200` is now a list where every compartment is stored separately; this can
be easily coerced into a `data.frame` for further plotting:

```{r}
res_200 <- cbind.data.frame(t = res_200[[1]], res_200[-1])
dim(res_200)
res_200[1:10, 1:4]
```

Let us define colors for our plot:
```{r }
## function to make colors transparent, stolen from adegenet::transp
transp <- function (col, alpha = 0.5) {
    res <- apply(col2rgb(col), 2,
                 function(c) rgb(c[1]/255, c[2]/255, 
                                 c[3]/255, alpha))
    return(res)
}

col <- rep(transp(sir_col, .1), each = 200)
```


```{r }
par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
matplot(res_200[, 1], res_200[, -1], xlab = "Time", ylab = "Number of individuals",
        type = "l", col = col, lty = 1)
legend("topright", lwd = 1, col = sir_col, legend = c("S", "I", "R"), bty = "n")

```


## A handy wrapper for replicates

The steps illustrated in the previous section can be wrapped into a simple
function:

```{r }
## x: instance of odin model
## t: time steps
## n: number of replicates
run_model <- function(x, t = 0:100, n = 1, ...) {
    res <- x$run(t, replicate = n, ...)
    res <- x$transform_variables(res)
    res <- cbind.data.frame(t = res[[1]], res[-1])
    attr(res, "n_compartments") <- length(x$names) - 1
    attr(res, "n_replicates") <- n
    attr(res, "compartments") <- x$names[-1]
    class(res) <- c("pretty_odin", class(res))
    res
}

sir_pal <- colorRampPalette(sir_col)

plot.pretty_odin <- function(x, pal = sir_pal, ...) {
    ## handle colors
    n_compartments <- attr(x, "n_compartments")
    n_replicates <- attr(x, "n_replicates")
    col_leg <- pal(n_compartments)
    alpha <- max(10 / n_replicates, 0.01)
    col <- rep(transp(col_leg, alpha), each = n_replicates)

    ## make plot
    par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
    matplot(x[, 1], x[, -1], xlab = "Time", ylab = "Number of individuals",
            type = "l", col = col, lty = 1, ...)
    legend("topright", lwd = 1, col = col_leg, bty = "n",
           legend = attr(x, "compartments"))
}

```

Here are two examples with 20 time steps, 50 simulations:
```{r }
plot(run_model(x, t = 0:20, n = 50))
```

and another one using 100 timesteps and 500 replicates:
```{r }
plot(run_model(x, t = 0:100, n = 500))
```



